---
title: "Exercise1"
author: "Michele Luca Puzzo, Marco Muscas, Shokoufeh Mansourihafshejani"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.1
Firstly we want to show how the algorithm computes the vector **y** at the first step. It is initialized as vector of zeros, it is p-dimensional. 
$$ \textbf{y}= (y_1, ..., y_i, ..., y_p) \hspace{0.2cm}$$
We are computing the i-th element of the vector **y** at first step. We observe that the frequency vector in the first step is made by all zeros except that for the j-th element that is one.
So it comes up that i-th element of y is equal to:
$$ y^{(1)}_i= \sum_{r=1}^p L_{ir} \cdot x^{(1)}_r = L_{ij} \cdot x^{(1)}_j = L_{ij} \cdot 1 = L_{ij}, \hspace{0.2cm} i\in \{1,...p\} $$
So at the end of the first step **y** will be equal to the column j-th of the random matrix: 
$$ \mathbf{y^{(1)}}= \mathbf{L_{:j}} $$

At time step k-1, j-th coordinate of **x^(k−2)^** increases by 1 to get **x^(k-1)^**. We call **f** the update of the frequency vector, made by all zeros except for the position j-th. We can break down the frequency vector at the step k-1 as sum of the frequency vector at step k-2 plus the update **f**
We know that the matrix multiplication has the distributive function respect to the sum so we can write:
$$ \mathbf{y^{k-1}} = L \cdot \mathbf{x^{k-1}} = L \cdot (\mathbf{x^{k-2}} + \mathbf{f}) = L \cdot \mathbf{x^{k-2}} + L \cdot  \mathbf{f} =  \mathbf{y^{k-2}} + \mathbf{L_{:j}}$$
The first addend was for the definition the vector **y** at the step k-1 while the second addend is how the column j-th of the matrix L: we have shown this computing the first step. 

## 1.2


```{r}
#initialization of the parameters. For now at random. Then we have to make some trials with bunch of them 
p = 100
d = 1000000
n = 10000
eps = 0.1
M <- 1000
```

I have to compute the frequency vector to compute its norm to check the *Johnson-Lindenstrauss lemma* but in the randomized algorithm we will not use it. We have decided to use the same raw stream because the probability in *JL lemma* simply accounts for the uncertainty implied by the randomness of L. 
We initialize **y** as a p-dimensional vector of zeros. 

```{r}

x <- vector(mode = "double", length = d)
D_n <- sample(1:d, n, replace = T)
for (i in D_n){
  x[i] <- x[i] + 1 
}
norm_x <- norm(x, type ="2")
```




```{r}

norm_y <- vector(mode = "double", length = M)
probab <- 0
for (r in 1:M){
  #compute matrix L, one for each simulation 
  L <- matrix(rnorm(p*d, 0, sqrt(1/p)), nrow = p, ncol = d)
  y <- vector(mode = "double", length = p)
  for (k in 1:n){
    #update the vector y
    y <- y + L[,D_n[k]]
  }
  norm_y[r] <- norm(y, type = "2")
  if ((norm_y[r]> (1-eps) * norm_x) & (norm_y[r] < (1+eps) * norm_x)){
  probab <- probab + 1 
  }
}
print(probab/M)
print(1 - exp(-p*eps^2))

```

```{r}

```
